---
title: "Project memo"
author: "Aronah, Clastel, Blake"
output: github_document
---

Summary: Agricultural practices have been discovered to have adverse effects on soil health and the quality of soils. Since the Industrial Revolution, humans have developed practices that may harm soil health and its essential properties. This project was based on a data set of soil qualities and properties from 218 different farmlands across Wisconsin and Minnesota from 2015 to 2017. The dataset explores how soil properties (e.g., texture, pH, and drainage class), as well as soil organic matter (SOM), are affected by specific farming and agricultural practices. Practices used among these farms include rotation of crops, tillage type and frequency (e.g., no-till, conventional till), use of cover crops, and manure application (e.g., timing, type, and frequency). We found that certain livestock have more effects on soil qualities like pH and that the amount of soil tilling can heavily impact the health and presence of water drainage in soil. 

Methods: The soil samples collected were obtained from six distinct regions of farmland and watersheds across the East of Wisconsin and Southeast of Minnesota. Soil samples were taken from areas identified by the farmer as areas with average crop yield and then frozen and stored (on the same day as the sample was collected) for 30 days to kill any microbes and stagnate organic matter mineralization. Samples were then thawed and analyzed for soil properties under the USDA NRCS Web Soil48 at the University of Wisconsin Soil and Forage Analysis.
Laboratory (Marshfield, Wisconsin). Soil organic matter (SOM), total organic carbon (TOC), total Nitrogen (TN), and total Carbon (TC) were determined through different tests, including heating and combustion as well as effervescence. Analysis of variance (ANOVA) tests were done to determine the individual effects of SOM, TOC, and TN. These samples were collected, tested, and analyzed to discover how agricultural practices and preexisting soil compositions influence SOM. 

Analysis: For our analysis we created a violin plot depicting variation in pH of soil in relation to the type of manure used on the soil. In addition the chart has a variable depicting the state of the manure whether it's solid, liquid, etc. In this chart it shows the variance of each manure type's impact on pH. The reason for using a violin plot was to see the density of the variation of ph.

For the second part of our analysis, we created a stacked bar plot  that counted the number of fields, the number of seasonal till passes, and evaluated the correlation with the soil’s drainage class (evaluated on a level 1-7, 7 retaining the least amount of water.) We wanted to use a bar plot to visualize percentages and demonstrate the correlation in a clearly interpretable graph. 

Conclusions: 
In terms of conclusions it’s hard to say whether the manure is the sole impact on pH so therefore it’s hard to make a final conclusion, however, within the visualization it’s pretty clear which are more consistent. In this case, Cows manure was the most consistent at staying in the ideal pH range for crop growth being between 6.8 and 7.5. More specifically the dairy cows were more consistent than beef cows due to the different diets.  

Tillage is the practice of turning the soil to allow for better airflow and make it easier for water to permeate the top layers of soil. However, tilling practices can contribute to soil loss and erosion, can reduce nutrients, and it disrupts the soil structure (Wallander 2020). Our data shows that the vast majority of well-drained soils were on farms with no-till or low-till practices. No-till or low-till practices allow soil structures to remain intact and avoid compaction, which is shown to directly correlate with better drainage.

### Load Packages
```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readr)
library(dplyr)
library(janitor)
library(ggthemes)
library(gapminder)
```

### Load data
```{r}
data2 <- read_csv("../data/Data2.csv")
```

### data cleaning Steps for Overall Data

### data cleaning step 1

```{r Data overveiw}
Data2_clean <- janitor::clean_names(data2)
glimpse(Data2_clean)
```

### Step 2:Recoding dairy/beef cow
```{r data-cleaning-step1}
Data2_clean <- Data2_clean |>
  mutate(
    mnr_source = case_when(
      mnr_source == "dairy" ~ "dairy cow",
      mnr_source == "beef" ~ "beef cow",
      TRUE ~ mnr_source
    )
  )
```

### Step 3: reordering/filtering missing factors/titles
```{r data-cleaning-step2}
Data2_clean |>
  mutate(mnr_solid = str_to_upper(mnr_solid)) |>
  filter(mnr_source != ".") |>
  filter(mnr_solid != ".") |>
  mutate(mnr_source = fct_relevel(
    mnr_source,
    "none",
    "beef cow",
    "dairy cow",
    "chicken",
    "swine",
    "turkey"
  ))
```

### Plot 1: pH levels of soil with differnt manure sources

#### Data cleanup steps specific to plot 1

```{r manure pH, fig.alt= "A vertical violin plot depicting the correlation of the use of different manure types with the pH of the soil. The types of manure being none, beef cow, dairy cow, chicken, swine, and turkey. In the chart the dairy cow has the most consistent pH that falls within the ideal range for plant growth depicted with a dotted line at 6.8 and 7.5."}
ggplot(Data2_clean, mapping = aes(x = mnr_source, y = p_h, fill = mnr_solid)) +
  geom_violin() +
  
  scale_fill_manual(
    values = c(
      "S" = "#404688FF",
      "L" = "#21908CFF",
      "B" = "#7AD151FF",
      "NONE" = "#FDE725FF"
    ),
    breaks = c("S", "L", "B", "NONE"),
    labels = c("Solid", "Liquid", "Both", "None")
  ) +
  
  
  
  
  labs(title = "pH levels of soil with different manure sources",
       x = "Manure Type",
       y = "pH",
       fill = "State of Manure") +
  geom_hline(aes(yintercept = 6.8), linetype = "dotted", size = 0.7) +
  geom_hline(aes(yintercept = 7.5), linetype = "dotted", size = 0.7) +
  scale_color_manual(name = "pH",
                     values = c("Unhealthy" = "black", "Healthy" = "black")) +
  geom_text(aes(x = "turkey", y = 7.6), label = "Ideal pH Range")


```
### Plot 2: Drainage and tilling correlation 
```{r water drainage correlation to tilling of the soil, fig.alt= "a horizontal violin plot is showing the correlation of tillage amount and frequencey compared to the amount of water drainage on a scale of 1-7: 1 being Very poorly drained and 7 being Excessively drained"}
#data cleaning for plot specifically
Data2_clean|> filter( springfall_till_passes != ".") |>
select(c(springfall_till_passes, drainage))|> 
  
#plot
ggplot( mapping = aes(x = drainage, y = springfall_till_passes))+
         geom_violin(fill = "darkgreen")+
  #scale_colour_viridis_d() + 
  labs(
  title = "Drainage and Tilling Correlation", subtitle = "Drainage class is rated on a scale from 1-7, ranging from 'very poorly drained' to  'excessively drained'. Tilling is the act of aerating the soil (typically with heavy machinery) to improve air circulation and water permeation. ", 
  x = "Drainage Class",
  y = "Spring- Fall Till Passes"
  
) 
```


### Plot 3: tillage impact on drainage

```{r tillage drainage, fig.alt = "A stacked horizontal bar chart depicting the correlation of till passes with the drainage of the field. On the y axis being number of till passes on the x axis the number of fields and the fill of the bars was how good the drainage was on the field."}
#data cleaning changed variable 
Data2_clean|> filter(springfall_till_passes != ".") |>  mutate(drainage = as.character(drainage)) |> group_by (springfall_till_passes, drainage) |> arrange(drainage)|>
ggplot(aes(fill = drainage, y = springfall_till_passes)) + 
  labs(title = "Tillage impact on drainage", x = "Number of fields", y = "Number of till passes", fill = "Drainage")+

  
  scale_fill_manual(
    values = c(
      "1" = "#440154FF",
      "2" = "#443A83FF",
      "3" = "#3B528BFF",
      "4" = "#2C728EFF",
      "5" = "#24868EFF",
      "6" = "#1F9A8AFF",
      "7" = "#47C16EFF"
    ),
    breaks = c("1", "2", "3", "4", "5", "6", "7"),
    labels = c(
      "Very poorly drained",
      "Poorly drained",
      "Somewhat poorly drained",
      "Moderately well drained",
      "Well drained",
      "Somewhat excessively well drained",
      "Excessively drained"
    )
  ) +
  

  geom_bar(position = "stack")


```
### Plot 4: bar plot of tilling and drainage

```{r}
Data2_clean|> group_by (springfall_till_passes, drainage) |> arrange(drainage)|>
ggplot(aes(fill = drainage, y = springfall_till_passes, x = drainage)) + 
  
    geom_bar(position = "stack", stat = "identity")
```
### Plot 5: Sand percentage Impact on Drainage

```{r fig.alt= "A line graph depicting the relationship between a soil's sand content and the drainage class. On the x axis, sand content is shown from 0-85%. On the y axis, drainage classes from 1-6 are shown."}
Data2_clean|>
  ggplot(aes(x = sand, y = drainage))+
  geom_smooth(color = "#440154FF", span = .5 ) + 
  labs( title = "Sand Percentage Impact on Drainage", x = "Sand (%)", y = "Drainage Class")
```
### Plot 6: Clay Percentage Impact on Drainage

```{r Clay Percentage Impact on Drainage, fig.alt= "A line graph depicting the relationship between a soil's clay content and the drainage class. On the x axis, clay content is shown from 0-46%. On the y axis, drainage classes from 1-6 are shown."}
Data2_clean|>
  ggplot(aes(x = clay, y = drainage))+
  geom_smooth(color = "#2C728EFF", span = .5) +
labs(title = "Clay Percentage Impact on Drainage", x = "Clay (%)" ,y = "Drainage Class")
```


## 4. Data Ethics Review

The data ethics review section will be introduced in a separate class and is not part of the original proposal deadline.
